---
title: "Vizualization"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

1. rnoaa = allows you to access public data
2. lines 15 - 19 --> bring data
3. 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Make Scatterlots!

Create my first scatterplot:
1. ggplot then type dataframe, define x & y
2. + and specify which type of graph you want
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + #aes = aesthetics, define x and y
  geom_point() #geom_point = scatterplot
```

New approach, same plot:
1. ggplot but use piping
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save & edit a plot object:
1. create a plot named "weather_plot"
```{r}
weather_plot =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot:

Start w/ the same one but make it fancy:

1. adding color option 
2. geom_smooth = add smooth curves that run through data points

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

What about the `aes` placement?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```

Let's facet smthings!

This separates graph via whatever. Here, we are separating graphs by name!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) + #alpha defines size of the points
  geom_smooth(se = FALSE, size = 2) + #size = thickness of the line
  facet_grid(. ~ name) #row = nothing, columns = name; separate graphs via name
```

Let's combine elements & do new plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## small notes

How many geoms have to exist? You can have how many ever geoms you want!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d()
```

## univariate plots 

### HISTOGRAMS!!!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color ...

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, color = name)) + #or replace "color = name" with "fill = name"
  geom_histogram(position = "dodge")
```

Let's try a new geometry!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```

### Boxplot!

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) + # x = name will result in 3 different plots per name
  geom_boxplot()
```

### Trendy plots :-)

1. to see median, etc, use stat_summary
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin() +
  stat_summary()
```

2. most popular plot of 2017 -- ridge plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Saving and Embedding

Let's save a scatterplot
```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5) 
```

What about embedding?
```{r, fig.width = 12, fig.height = 6}
weather_plot
```

```{r, fig.width = 6, fig.height = 3}
weather_plot
```
